---
title: "General Social Survey"
author: "Grace carter"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

mydata5 <- filter(mydata4, CAPPUN=="FAVOR" | CAPPUN=="OPPOSE")

```




# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.

We are comparing what happened with the people who voted in 2012 vs 2016. Do those who vote tend to always vote or does the election influence it? Did those who voted in 2012, still vote 2016 or is there a difference? Of those who did not vote in 2012, did they not vote in 2016? There is a noticable consistency of voters who didn't vote in 2012, not voting in 2016. But the voters have a larger shift from voting 2012 to not voting in 2016. Are these independent or are they related?   

Null hypothesis: Voting one year to the next is not related.
Alternative hypothesis, they are related.
If P is < .05 reject null hypothesis.
If P > .05 fail to reject

How are the "voted in 2016" and "voted in 2012" related?  

$H_0:$ voting in 2016 and 2012 are unrelated  
$H_a:$ voting in 2016 and 2012 are related  

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use categorical variables, and a Chi-squared test to test the hypothesis that voting in 2016 is / is not independent of voting in 2012. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the voting trends between 2012 and 2016

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent", xlab= 'voted in 2016') 

```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation:  

```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

$N_0$: Voting in 2016 is independent of voting in 2012.
$N_n$: Voting in 2016 is dependent on voting in 2012.

The hypothesis was that voting in 2012 is related to voting in 2016.    
If you voted in 2012 it is likely you voted in 2016.  We did a chi-squared test to see if voting in 2016 was dependent on voting in 2012. Result was 75.75
The P-value was < 2.2e-16 which is smaller then .05, so we reject the null hypothesis.  
Because the response and explanatory values are catagorical, the bar plot and chi-squared test test for independence. 

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

Information came from general social survey 2018. We want to know if there is a correlation between belief in the BigBang and Belief in a Postlife. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use categorical variables, and a Chi-squared test to test the hypothesis that belief int he Bigband is / is not independent of belief in postlife. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the data on belief in BigBand and Postlife.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent", xlab= 'BELIEVE IN POSTLIFE', main='Believe in BigBang') 
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~BIGBANG+POSTLIFE, data=mydata3)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
votetab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(votetab)
```

$H_0$: belief in bigbang is independent of belief in postlife.
$H_a$: belief in bigbang is dependent on belief in postlife.

The null hypothesis was that belief in postlife is unrelated to belief in bigbang.   

There is a correlation between not believing in postlife and believing in bigbang. There is a strong correlation between not believeing in post-life and believing in bigbang. However, there is a larger percent of people who believe in both bigbang and post life. There are very few who do not believe in either bigbang nor postlife.

If you do not believe in postlife, it is likely you do believe in bigbang. We did a chi-squared test to see if postlife was dependent on bigbang, the result was 7.4181
The P-value was = 0.004074 which is smaller then .05, so we reject the null hypothesis.  
Because the response and explanatory values are catagorical, the bar plot and chi-squared test test for independence.  

  
# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

Information came from general social survey 2018.Is there a relationship between an opinion on the death penatly (CAPPUN) and the opinion on courts (COURTS).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use categorical variables, and a Chi-squared test to test the hypothesis that the opinion on cappun is / is not independent of opinion of courts. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the data on opinon on CAPPUN and opinion on courts.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent", xlab= 'View on Capital Punishment', main='View on Courts')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~CAPPUN+COURTS, data=mydata5)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
votetab=xtabs(~CAPPUN+COURTS, data=mydata5)
fisher.test(votetab)
```


$H_0$: There is no relationship between the view on the court and the view on capital punishment. 
$H_a$: There is a relationship between he view on the court and the view on capital punishment. 

The p-value = 4.35e-10 which is smaller than .05, so we reject the null hypothesis.  
Because the response and explanatory values are catagorical, the bar plot and chi-squared test test for independence. 
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

All three examples seemed to have some correlation in their variables. The 2012 voters mostly voted in 2016, and non's didn't; it seemed more that people tended to vote or not consistently. the non-belief in post life seemed directly related to belief in the big bang. Though there was a larger group who believed in both the big bang and postlife. There was a small group who believed in neither. The capital punishment went directly with the group who felt the courts were too lenient. 
